Veasel meets FOX: an ensemble approach to entity linking based on vector similarity

### 2015-07-10 paper.tex
minor changes

%opening
\title{Veasel meets Fox: an ensemble approach to entity linking based on vector similarity }
\titlerunning{Veasel meets Fox}

\author{Felix Tristram \and Christina Unger \and Sebastian Walter \and Philipp Cimiano}
\authorrunning{Felix Tristram \and Christina Unger \and Sebastian Walter \and Philipp Cimiano}
% (feature abused for this document to repeat the title also on left hand pages)

@@

\begin{abstract}
The task of entity linking consists in disambiguating named entities occurring in textual data by linking them to an identifier in a knowledge base that represents the real-world entity they denote. We present \emph{Veasel}, a novel approach that is based on an ensemble of different features that is induced using a Support Vector Machine. We compare our approach to state-of-the-art tools such as Fox and DBpedia spotlight, showing that it outperforms both on the AIDA/CoNLL dataset and provides comparable results on the KORE50 dataset. An important feature of Veasel is its speed.

@@

The task of entity linking consists in disambiguating named entities occurring in textual data by linking them to an identifier in a knowledge base that represents the real-world entity they denote. We present \emph{Veasel}, a novel approach that is based on an ensemble of different features that is induced using a Support Vector Machine.  Given a named entity and a resource candidate, Veasel combines four features that quantify i) the likelihood that the actual name refers to the resource candidate (extracted from Wikipedia anchors), ii) the similarity of the textual context in which the name occurs to the textual context of the resource candidate (measures as the bag-of-words similarity using cosine), iii) the overlap between the semantic neighborhood signature of the resource candidate and the named entities mentioned in the larger context of the named entity, and iv) the prominence of the resource candidate as measured by its page rank in the DBpedia graph. An optimal linear model is trained using an SVM to induce an ensemble that combines these four features into an overall decision whether the named entity refers to a candidate resources or not.

@@

We compare our approach to state-of-the-art tools such as Fox and DBpedia spotlight, showing that it outperforms both on the AIDA/CoNLL dataset and provides comparable results on the KORE50 dataset. Further, we show that Veasel beats state-of-the-art named entity linking approaches in terms of speed.

In the following section we describe the approach, and then report evaluation results in Section~\ref{sec:results}.

@@

The input data for {\acronym} is a regular piece of text, for example a newspaper article, in which entity-representing sequences of tokens (\emph{fragments}) are already identified (possibly by a named entity recognition tool). For a given fragment, the algorithm's task is to determine the matching real-world entity. It does so by picking the most likely entity from a set of entity candidates, e.g. specified in a knowledge base, in our case Wikipedia.

A fragment which represents a real-world entity is called an \textit{anchor}. To gather a database of anchors and their corresponding entities, we extracted all anchor texts from a Wikipedia dump\footnote{\url{http://en.wikipedia.org/wiki/Wikipedia:Database_download}}. Within a dump hyperlinks to other pages appear in the form \texttt{[[Barack Obama|President Obama]]}, 
where {\texttt{Barack Obama} refers to an entity\footnote{\url{http://en.wikipedia.org/wiki/Barack_Obama}} and \texttt{President Obama} is the anchor string which appears in the page's text. A map is generated by parsing all Wikipedia articles (in a chosen language) and counting the occurrences of anchor-entity pairs. The candidates for a given fragment are now found by looking up all entities that were referred to by the fragment within Wikipedia.

@@

The semantic neighborhood signature for an entity is computed in the same ways as Babelfy \cite{babelfy} does, except that instead of the BabelNet\footnote{\url{http://babelnet.org/}} knowledge base we use a graph of interlinked Wikipedia pages, where each page is a vertex and each link between pages is an edge. The idea is to identify for each resource all other resources it is closely related to. This is a two step process. First, the edges are assigned weights so that edges in denser parts of the graph become more important. The weight of an edge $\mathrm{weight}(v, v')$ is the number of directed triangles (directed cycles of length 3) it appears in:
$$\mathrm{weight}(v, v') := |\{(v, v', v'') : (v, v'), (v', v''), (v'', v) \in E\}| + 1$$

@@

In its essence we thus compute a measure similar to the \emph{personalized page rank} \cite{personalized_page_rank} using a random walk approach similar to the one implemented in Babelfy. This step is performed offline.

@@

Therefore a candidate's semantic neighborhood signature similarity score increases as entities from their semantic neighborhood signature appear more often as candidates in a given input.
The top 10 elements in the semantic neighborhood signature of Barack Obama are for instance:
\begin{enumerate}
\item United\_States\_House\_of\_Representatives\_elections,\_2010
\item Barack\_Obama
\item Presidency\_of\_Barack\_Obama
\item United\_States\_presidential\_election,\_2008
\item Barack\_Obama\_presidential\_primary\_campaign,\_2008
\item First\_inauguration\_of\_Barack\_Obama
\item Democratic\_Party\_(United\_States)
\item George\_W.\_Bush
\item Republican\_Party\_(United\_States)
\item Joe\_Biden
\end{enumerate}

@@

One of the biggest strengths of {\acronym} is its speed. 
Running it on the full datasets took less than one minute for KORE50 and about 45 seconds for AIDA.
\todo[inline]{Redo measurements with the latest approach.}
%As comparison, running Babelfy on the KORE50 dataset takes about 15 minutes, and running it on AIDA had to be stopped after 164 hours.

@@

We have presented Veasel, a new entity linking approach that exploits an ensemble of features that are linearly combined using a Support Vector Machine. We have shown that the approach works very well when the contexts are longer, clearly outperforming Fox and DBpedia Spotlight on the AIDA CoNLL dataset. On the KORE50 dataset where contexts are shorter, it achieves comparable results to these approaches. In future work, we will release a service that can be publicly accessed using a similar interface to DBpedia Spotlight and Fox. We will also extend the approach to other languages and perform a more systematic evaluation using the GERBIL framework \cite{gerbil}.

>>>>

%opening
\title{Veasel meets FOX: an ensemble approach to entity linking based on vector similarity }
\titlerunning{Veasel meets FOX}

\author{Felix Tristram \and Sebastian Walter \and Philipp Cimiano  \and Christina Unger}
\authorrunning{Felix Tristram \and Sebastian Walter \and Philipp Cimiano  \and Christina Unger}
% (feature abused for this document to repeat the title also on left hand pages)

@@

\begin{abstract}
The task of entity linking consists in disambiguating named entities occurring in textual data by linking them to an identifier in a knowledge base that represents the real-world entity they denote. We present \emph{Veasel}, a novel approach that is based on an ensemble of different features that is trained using a Support Vector Machine. We compare our approach to state-of-the-art tools such as FOX and DBpedia spotlight, showing that it outperforms both on the AIDA/CoNLL dataset and provides comparable results on the KORE50 dataset. An important feature of Veasel is its speed.

@@

The task of entity linking consists in disambiguating named entities occurring in textual data by linking them to an identifier in a knowledge base that represents the real-world entity they denote. We present \emph{Veasel}, a novel approach that is based on an ensemble of different features that is trained using a Support Vector Machine.  Given a named entity and a resource candidate, Veasel combines four features that quantify i) the likelihood that the actual name refers to the resource candidate (extracted from Wikipedia anchors), ii) the similarity of the textual context in which the name occurs to the textual context of the resource candidate (measured as the bag-of-words similarity using cosine), iii) the overlap between the semantic neighborhood signature of the resource candidate and the named entities mentioned in the larger context of the named entity, and iv) the prominence of the resource candidate as measured by its page rank in the DBpedia graph. An optimal linear model is trained using an SVM to induce an ensemble that combines these four features into an overall decision whether the named entity refers to a candidate resource or not.

@@

We compare our approach to state-of-the-art tools such as FOX\cite{foxdemo} and DBpedia spotlight\cite{spotlight}, showing that it outperforms both on the AIDA/CoNLL dataset and provides comparable results on the KORE50 dataset. Further, we show that Veasel beats state-of-the-art named entity linking approaches in terms of speed.

In the following section we describe the approach and then report evaluation results in Section~\ref{sec:results}.

@@

The input data for {\acronym} is a regular piece of text, for example a newspaper article, in which entity-representing sequences of tokens (\emph{fragments}) are already identified (we use the Stanford NER \cite{ner}). For a given fragment, the algorithm's task is to determine the matching real-world entity. It does so by picking the most likely entity from a set of entity candidates, e.g. specified in a knowledge base, in our case DBpedia.

A fragment which represents a real-world entity is called an \textit{anchor}. To gather a database of anchors and their corresponding entities, we extracted all anchor texts from a Wikipedia dump\footnote{\url{http://en.wikipedia.org/wiki/Wikipedia:Database_download}}. Within a dump, hyperlinks to other pages appear in the form \texttt{[[Barack Obama|President Obama]]}, 
where {\texttt{Barack Obama} refers to an entity\footnote{\url{http://en.wikipedia.org/wiki/Barack_Obama}} and \texttt{President Obama} is the anchor string which appears in the page's text. A map is generated by parsing all Wikipedia articles (in a chosen language) and counting the occurrences of anchor-entity pairs. The candidates for a given fragment are found by looking up all entities that were referred to by the fragment within Wikipedia.

@@

The semantic neighborhood signature for an entity is computed in a similar way as Babelfy \cite{babelfy} does, except that instead of the BabelNet\footnote{\url{http://babelnet.org/}} knowledge base we use a graph of interlinked Wikipedia pages, where each page is a vertex and each link between pages is an edge. The idea is to identify for each resource all other resources it is closely related to. This is a two step process. First, the edges are assigned weights so that edges in denser parts of the graph become more important. The weight of an edge $\mathrm{weight}(v, v')$ is the number of directed triangles (directed cycles of length 3) it appears in:
$$\mathrm{weight}(v, v') := |\{(v, v', v'') : (v, v'), (v', v''), (v'', v) \in E\}| + 1$$

@@

We compute a measure similar to the \emph{personalized page rank} \cite{personalized_page_rank} using a random walk approach similar to the one implemented in Babelfy. This step is performed offline.

@@

Therefore a candidate's semantic neighborhood signature similarity score increases as entities from their semantic neighborhood signature appear more often as candidates in a given input.
%The top 10 elements in the semantic neighborhood signature of Barack Obama are for instance:
%\begin{enumerate}
%\item United\_States\_House\_of\_Representatives\_elections,\_2010
%\item Barack\_Obama
%\item Presidency\_of\_Barack\_Obama
%\item United\_States\_presidential\_election,\_2008
%\item Barack\_Obama\_presidential\_primary\_campaign,\_2008
%\item First\_inauguration\_of\_Barack\_Obama
%\item Democratic\_Party\_(United\_States)
%\item George\_W.\_Bush
%\item Republican\_Party\_(United\_States)
%\item Joe\_Biden
%\end{enumerate}

@@

One of the biggest strengths of {\acronym} is its speed. 
Running it on the full datasets took less than one minute for KORE50 and about 45 seconds for AIDA.
%\todo[inline]{Redo measurements with the latest approach.}
%As comparison, running Babelfy on the KORE50 dataset takes about 15 minutes, and running it on AIDA had to be stopped after 164 hours.

@@

We have presented Veasel, a new entity linking approach that exploits an ensemble of features that are linearly combined using a Support Vector Machine. We have shown that the approach works very well when the contexts are longer, clearly outperforming FOX and DBpedia Spotlight on the AIDA CoNLL dataset. On the KORE50 dataset where contexts are shorter, it achieves comparable results to these approaches. In future work, we will release a service that can be publicly accessed using a similar interface to DBpedia Spotlight and FOX. We will also extend the approach to other languages and perform a more systematic evaluation using the GERBIL framework \cite{gerbil}.
